---
title: "BiasExamples"
author: "David Currie"
date: "16/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## ICES Special Request - Example of functions to highlight sources of potential bias

```{r loadFunctions, message=FALSE, warning=FALSE}
# Load our functions
source("RDBES_Bias_Functions.R")
library(kableExtra)
```

```{r loadData, echo=FALSE}

# IMPORTANT: Hack to stop write.csv changing numbers to scientific notation
options(scipen=500) # big number of digits

## STEP 1) LOAD OUR DATA

# Load the validation data
validationData <- getValidationData(downloadFromGitHub = FALSE, fileLocation = './tableDefs/BaseTypes.xsd')
#validationData <- getValidationData(downloadFromGitHub = TRUE, fileLocation = './tableDefs/BaseTypes.xsd')

# 11/9/2020 Temp fix because the validation fields aren't up to date :-(
validationData[validationData$type == 'tRS_Stratification','type'] <- 'tYesNoFields'

# Load the reference data: either refresh from ICES or just use a local copy
allowedValues <- loadReferenceData(downloadFromICES = FALSE)
#allowedValues <- loadReferenceData(downloadFromICES = TRUE, validationData=validationData)

# Load the lists of tables required for each hierarchy: either refresh from ICES or just use a local copy
allRequiredTables <- getTablesInHierarchies(downloadFromGitHub = FALSE, fileLocation = './tableDefs/')
#allRequiredTables <- getTablesInHierarchies(downloadFromGitHub = TRUE, fileLocation = './tableDefs/')

# Load the RDBES data from the database - you can either write your own database connection string in a format similar to this: 'driver=SQL Server;server=mysqlhost;database=mydbname;trusted_connection=true' or just manually create a named list of data fames in the correct format
# IMPORTANT - if you are just going to use your own list of data frames make sure you don't have factors in them - my code assumes the data frames were created using stringsAsFactors = FALSE
myRDBESData <- loadRDBESData(readRDS("connectionString.RDS"))
```

## Selection methods

The first function highlights non-probabilistic seection methods, and the numebr of units that were intended to be sampled but weren't - these are both potentially sources of bias.

```{r selectionMethods}
myResult <- summariseSelectionMethods(hierarchyToCheck = 'H1', yearToUse = 2019, country = 'IE', rdbesData = myRDBESData, requiredTables = allRequiredTables)

# This is the data we'll print - after soem small formatting changes
dataToPrint <- myResult$overallSummary
# change NA to blanks
dataToPrint[is.na(dataToPrint)] <- ""
# get the "non caculated" column numbers - we'll format these differently
myNonCalcColumns <- which(names(dataToPrint) %in% names(dataToPrint)[grepl('^..numberNotSampledCalc$',names(dataToPrint))])

# Print the overall summary
dataToPrint %>% 
    mutate_all(~cell_spec(
        .x,
        background = ifelse(.x %in% c('NPAH','NPEJ','NPQS'), 'red','white'),
        color = ifelse(.x %in% c('NPAH','NPEJ','NPQS'), 'white','black'))) %>%
  kbl(escape = F) %>%
  column_spec(myNonCalcColumns, bold=T) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

# This is the data we'll print - after soem small formatting changes
dataToPrint <- myResult$summaryDataJoined
# change NA to blanks
dataToPrint[is.na(dataToPrint)] <- ""
# get the "non caculated" column numbers - we'll format these differently
myNonCalcColumns <- which(names(dataToPrint) %in% names(dataToPrint)[grepl('^..numberNotSampledCalc$',names(dataToPrint))])

# Print the more detailed infromation by stratum
dataToPrint %>% 
    mutate_all(~cell_spec(
        .x,
        background = ifelse(.x %in% c('NPAH','NPEJ','NPQS'), 'red','white'),
        color = ifelse(.x %in% c('NPAH','NPEJ','NPQS'), 'white','black'))) %>%
  kbl(escape = F) %>%
  column_spec(myNonCalcColumns, bold=T) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
